### 2024-08-29 paper.tex
more linalg

\begin{lemma}\label{gwr_irred_dim}
    \(G_{wr}\) is irreducible and has dimension \(\sum_{ij} \min(\lambda_i', \lambda_j')\), where \(\lambda' = (\lambda_1, ..., \lambda_{w - 1}, \lambda_w - r, \lambda_{w + 1}, ..., \lambda_k)\).
\end{lemma}
\begin{proof}
    The closure of \(G_{wr}\) in \(\gl_m\) is \(V = \{Y \in \mathfrak{z}_{\gl_m}(X) : e_{wr} Y = e_{wr}\}\).
    In the case that \(r = 0\), the constraint that \(e_{wr}Y = e_{wr}\) is no constraint at all, so we have \(G_{wr} = G\), and the result follows from \cref{g_irred_dim}.
    \par In the case that \(r > 0\), we appeal to \cref{nilpotent_centralizer}, which tells us exactly what \(\mathfrak{z}_{\gl_m}(X)\) looks like.
    Note that \(V = \{Y + I : Y \in \mathfrak{z}_{\gl_m}(X), e_{wr}Y = 0\}\).
    The constraint \(e_{wr}Y = 0\) is just saying that a certain row of \(Y\) must be all zeroes.
    Looking at the form of \(\mathfrak{z}_{\gl_m}(X)\), we see that requiring this row equal to zero for \(M \in \mathfrak{z}_{\gl_m}(X)\) is just the statement that \(M_{wr}\) is of the form
    \[\]

@@

\begin{lemma}
    Let \(\lambda = (\lambda_1, ..., \lambda_k)\) be a partition of \(m\).
    Let \(w \in \{1, ..., k\}\) and \(r \in \{1, ..., \lambda_w\}\).
    Let \(i = [\sum_{j < w} \lambda_j] + (r - 1)\).
    The set of \(M \in \mathfrak{z}_{\gl_m}(J(\lambda))\) such that \(\mathbbm{1}_iM = \mathbbm{1}_i\) is  
\end{lemma}
Now we determine which matrices of this form satisfy \(aA = a\).
For simplicity, we will instead find the matrices \(A\) such that \(aA = 0\) (so that \(a(A + I) = a\)).
(Note that \(I \in C_1\), and \(C_1\) is closed under addition, so this is really the same thing.)
\par In the case that \(a = 0\), clearly every \(A\) works.
Otherwise, let \(i_0, j_0\) be such that \(a_{i_0,j_0} = 1\).
Now, clearly the constraint that \(aA = 0\) is just saying that the \((i_0,j_0)\)th row of \(A\) must be zero.
That is, for each \(j\) the \(j_0\)th row of \(A_{i_0j}\) must be zero.
This just requires that for each \(j\), we must have 
\[v_{ij} = \begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i - j_0, \lambda_j)}\) can be chosen freely.
So, we have now found the set 
\[C_2 := \{A \in \gl_m : AX = XA, aA = a\}.\]
It is the matrices of the form \(I + A\), where \(A_{ij}\) is a block matrix of size \(\lambda_i \times \lambda_j\) with 
\[A_{ij} = T(v_{ij}) = T\begin{pmatrix}
    0\\
    v_{ij}'
\end{pmatrix},\]
where \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i, \lambda_j)}\) in the case \(i \neq i_0\) and \(v_{ij}' \in \mathbb{C}^{\min(\lambda_i - j_0, \lambda_j)}\) if \(i = i_0\).

\subsection{A `normalization' fact about Jordan bases}

\begin{lemma}\label{normalization_helper_one}
    For any finite-dimensional \(V\), nilpotent \(A : V \to V\), and linear \(f : V \to \mathbb{C}\), there is a Jordan basis \(e_{ij}\) for \(A\) such that there is at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}) \neq 0\).
\end{lemma}
\begin{proof}
    For any Jordan basis \((e_{ij})_{ij}\) of \(A\), define
    \[S((e_{ij})_{1 \leq i \leq k, 1 \leq j \leq \lambda_i}) := \sum_i \begin{cases}-1, & \forall j. \; f (e_{ij}) = 0 \\ \lambda_i - \min\{j : f (e_{ij}) \neq 0\}, & \textrm{otherwise} \end{cases}.\]
    We proceed by induction on the measure \(S\).
    That is, let \((e_{ij})_{ij}\) be a Jordan basis for \(A\).
    Our inductive hypothesis is that if there exists a Jordan basis \((e_{ij}')_{ij}\) with \(S((e_{ij}')_{ij}) < S((e_{ij})_{ij})\), then we get the desired conclusion.
    \par Now, we have two cases.
    In the first case, \((e_{ij})_{ij}\) already satisfies the desired property.
    In this case we are done.
    In the other case, there exist \(i_1, j_1, i_2, j_2\) with \(i_1 \neq i_2\), and \(f(e_{i_1j_1}) \neq 0\), and \(f(e_{i_2j_2}) \neq 0\).
    We let \(j_1, j_2\) be minimal with this property, so that \(\forall j < j_1. \; f(e_{i_1j}) = 0\), and \(\forall j < j_2. \; f(e_{i_2j}) = 0\).
    Wlog, we assume that \(\lambda_{i_1} - j_1 \leq \lambda_{i_2} - j_2\).
    \par By our inductive hypothesis, all we need to do is find a Jordan basis \((e_{ij}')_{ij}\) with \(S((e_{ij}')_{ij}) < S((e_{ij})_{ij})\).
    This is what we do.
    Define \(e'_{ij}\) as follows.
    \begin{itemize}
        \item \(e_{i_1, \lambda_1}' := e_{i_1, \lambda_1} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}\)
        \item For \(j < \lambda_1\), \(e_{i_1, j}' := A^{\lambda_{i_1} - j} e_{i_1, \lambda_{i_1}}'\)
        \item For \(i \neq i_1\), \(e_{ij}' := e_{ij}\).
    \end{itemize}
    Clearly this is a Jordan basis for \(A\).
    Further, I claim that \(S((e'_{ij})_{ij})  < S((e_{ij})_{ij})\).
    It suffices to show that \(\forall j \leq j_1. \; f (e_{i_1,j}) = 0\).
    We have 
    \[f (e_{i_1,j}') = f\left(A^{\lambda_{i_1} - j}\left(e_{i_1, \lambda_1} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}\right)\right) =\]
    \[f\left(e_{i_1, j} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})} e_{i_2, j_2 + (j - j_1)}\right) = f(e_{i_1, j}) - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})} f(e_{i_2, j_2 + (j - j_1)}).\]
    Clearly (by design), this expression is zero when \(j = j_1\).
    And for \(j < j_1\), we have \(f(e_{i_1, j}) = f(e_{i_2, j_2 + (j - j_1)}) = 0\), so it is zero then as well.
    Hence the measure \(S\) of this new basis is smaller, as desired.
\end{proof}

>>>>

\begin{lemma}\label{gwr_irred_dim}
    \(G_{wr}\) is irreducible and has dimension \(\sum_{ij} \min(\lambda_i', \lambda_j')\), where \(\lambda' = (\lambda_1, ..., \lambda_{w - 1}, \lambda_w - r, \lambda_{w + 1}, ..., \lambda_k)\).
\end{lemma}
\begin{proof}
    The closure of \(G_{wr}\) in \(\gl_m\) is \(V = \{Y \in \mathfrak{z}_{\gl_m}(X) : e_{wr} Y = e_{wr}\}\).
    In the case that \(r = 0\), the constraint that \(e_{wr}Y = e_{wr}\) is no constraint at all, so we have \(G_{wr} = G\), and the result follows from \cref{g_irred_dim}.
    \par In the case that \(r > 0\), we observe that \(V = \{Y + I : Y \in \mathfrak{z}_{\gl_m}(X), e_{wr}Y = 0\}\).
    The constraint \(e_{wr}Y = 0\) is just saying that a certain row of \(Y\) must be all zeroes.
    Looking at \cref{nilpotent_centralizer_row_zero}
    Looking at the form of \(\mathfrak{z}_{\gl_m}(X)\), we see that requiring this row equal to zero for \(M \in \mathfrak{z}_{\gl_m}(X)\) is just the statement that \(M_{wr}\) is of the form
    \[\]

@@

\begin{corollary}\label{nilpotent_centralizer_row_zero}
    Let \(\lambda = (\lambda_1, ..., \lambda_k)\) be a partition of \(m\).
    Let \(w \in \{1, ..., k\}\) and \(r \in \{1, ..., \lambda_w\}\).
    Let \(i = [\sum_{j < w} \lambda_j] + r\).
    The set of \(M \in \mathfrak{z}_{\gl_m}(J(\lambda))\) such that the \(i\)th row of \(M\) is equal to zero is the set of matrices
    \[M = \begin{pmatrix}
        M_{11} & \cdots & M_{1k} \\
        \vdots & & \vdots\\
        M_{k1} & \cdots & M_{kk}
    \end{pmatrix},\]
    where each \(M_{ij}\) is a \(\lambda_i \times \lambda_j\) matrix, such that:
    \begin{itemize}
        \item For \(i \neq w\), \(M_{ij}\) is lower-left Toeplitz
        \item Each \(M_{wj}\) is of the form
        \[M_{wj} = \begin{pmatrix}
            0\\
            M_{wj}'
        \end{pmatrix},\]
        where \(M_{wj}'\) is a \((\lambda_w - r) \times \lambda_j\) lower-left Toeplitz matrix.
    \end{itemize}
\end{corollary}

\subsection{A `normalization' fact about Jordan bases}
\begin{lemma}\label{normalization_helper_one}
    For any finite-dimensional \(V\), nilpotent \(A : V \to V\), and linear \(f : V \to \mathbb{C}\), there is a Jordan basis \(e_{ij}\) for \(A\) such that there is at most one \(i\) such that there exists \(j\) such that \(f(e_{ij}) \neq 0\).
\end{lemma}
\begin{proof}
    For any Jordan basis \((e_{ij})_{ij}\) of \(A\), define
    \[S((e_{ij})_{1 \leq i \leq k, 1 \leq j \leq \lambda_i}) := \sum_i \begin{cases}-1, & \forall j. \; f (e_{ij}) = 0 \\ \lambda_i - \min\{j : f (e_{ij}) \neq 0\}, & \textrm{otherwise} \end{cases}.\]
    We proceed by induction on the measure \(S\).
    That is, let \((e_{ij})_{ij}\) be a Jordan basis for \(A\).
    Our inductive hypothesis is that if there exists a Jordan basis \((e_{ij}')_{ij}\) with \(S((e_{ij}')_{ij}) < S((e_{ij})_{ij})\), then we get the desired conclusion.
    \par Now, we have two cases.
    In the first case, \((e_{ij})_{ij}\) already satisfies the desired property.
    In this case we are done.
    In the other case, there exist \(i_1, j_1, i_2, j_2\) with \(i_1 \neq i_2\), and \(f(e_{i_1j_1}) \neq 0\), and \(f(e_{i_2j_2}) \neq 0\).
    We let \(j_1, j_2\) be minimal with this property, so that \(\forall j < j_1. \; f(e_{i_1j}) = 0\), and \(\forall j < j_2. \; f(e_{i_2j}) = 0\).
    Wlog, we assume that \(\lambda_{i_1} - j_1 \leq \lambda_{i_2} - j_2\).
    \par By our inductive hypothesis, all we need to do is find a Jordan basis \((e_{ij}')_{ij}\) with \(S((e_{ij}')_{ij}) < S((e_{ij})_{ij})\).
    This is what we do.
    Define \(e'_{ij}\) as follows.
    \begin{itemize}
        \item \(e_{i_1, \lambda_1}' := e_{i_1, \lambda_1} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}\)
        \item For \(j < \lambda_1\), \(e_{i_1, j}' := A^{\lambda_{i_1} - j} e_{i_1, \lambda_{i_1}}'\)
        \item For \(i \neq i_1\), \(e_{ij}' := e_{ij}\).
    \end{itemize}
    Clearly this is a Jordan basis for \(A\).
    Further, I claim that \(S((e'_{ij})_{ij})  < S((e_{ij})_{ij})\).
    It suffices to show that \(\forall j \leq j_1. \; f (e_{i_1,j}) = 0\).
    We have 
    \[f (e_{i_1,j}') = f\left(A^{\lambda_{i_1} - j}\left(e_{i_1, \lambda_1} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})}e_{i_2, j_2 + (\lambda_{i_1} - j_1)}\right)\right) =\]
    \[f\left(e_{i_1, j} - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})} e_{i_2, j_2 + (j - j_1)}\right) = f(e_{i_1, j}) - \frac{f(e_{i_1, j_1})}{f(e_{i_2, j_2})} f(e_{i_2, j_2 + (j - j_1)}).\]
    Clearly (by design), this expression is zero when \(j = j_1\).
    And for \(j < j_1\), we have \(f(e_{i_1, j}) = f(e_{i_2, j_2 + (j - j_1)}) = 0\), so it is zero then as well.
    Hence the measure \(S\) of this new basis is smaller, as desired.
\end{proof}
